---
title: "Class Project"
author: "Owen Jetton"
date: "2/20/2021"
output: github_document
---

# Data

```{r setup}
knitr::opts_chunk$set(echo = TRUE)
 
# Download and load packages
library(pacman)

p_load(readr, dplyr, ggplot2, readxl, magrittr, janitor, utils, tidyverse, tidymodels,
       tune, glmnet, recipes)
```

```{r reading data}
# Read CO2 data from 2016 - 19
co2_data_2016 = read_xlsx("C:/Users/ojett/Documents/EC424/ec424project/emissions2016.xlsx",
                     sheet = 1) %>% mutate(year = 2016)
                        # add year variable which will be helpful for merging data later

co2_data_2017 = read_xlsx("C:/Users/ojett/Documents/EC424/ec424project/emissions2017.xlsx",
                     sheet = 1) %>% mutate(year = 2017)

co2_data_2018 = read_xlsx("C:/Users/ojett/Documents/EC424/ec424project/emissions2018.xlsx",
                     sheet = 1) %>% mutate(year = 2018)

co2_data_2019 = read_xlsx("C:/Users/ojett/Documents/EC424/ec424project/emissions2019.xlsx",
                     sheet = 1) %>% mutate(year = 2019)


# Upload heating and cooling days data
heat_cool_df = read.delim("https://www.ncei.noaa.gov/pub/orders/CDODiv6010178275591.txt",
                          sep = ",") 


```

# Cleaning

```{r Cleaning Data}

# Merging CO2 to get larger data set
data_full = do.call("rbind", list(co2_data_2016, co2_data_2017,
                                  co2_data_2018, co2_data_2019))


# Cleaning the full CO2 data
data_full %<>% clean_names() %>%
  
    # only care about power plants that generate electricity
  filter(generation_k_wh > 0) %>% 
  
    # remove irrelevent or repretitive variables
  select(-c(eia_balancing_authority_region, balancing_authority_name, nerc_region, 
            fuel_units, metric_tonnes_of_co2_emissions, plant_name))



# Cleaning the Heating and Cooling days data
heat_cool_df %<>% clean_names() %>% 
    # select relevant variables
  select(state_code, year_month, cdd, hdd) %>%
  
    # make year variable from year_month variable
  mutate(year = strtrim(year_month,4)) %>%
    
    # grouping by months, and taking the sum to get the total number
      # of heating and cooling days in a year
  group_by(state_code, year) %>%
  mutate(cooling_days = sum(cdd),
         heating_days = sum(hdd),
         state_code = factor(state_code)) %>%
  
    # selecting relevant 
  select(state_code, year, cooling_days, heating_days) %>%
  distinct()

# Need the state names to correspond with the right regions:
  # read dataset that has state names for corresponding regions
  states_region = read_xlsx("C:/Users/ojett/Documents/EC424/ec424project/region_to_state.xlsx")
  
  # add state names to respective region in 
heat_cool_df = merge(heat_cool_df, states_region, by = "state_code")


# Add heating and cooling data to main dataframe

data_full = merge(data_full, heat_cool_df, by = c("state", "year"))


# Renaming CO2 emissions variable to "emissions"
  # I'm doing this because it is an important variable and will need to be written a lot
data_full %<>% rename(emissions = tons_of_co2_emissions) %>%
  
  # If there is no balancing authority
  mutate(balancing_authority_code = if_else(is.na(balancing_authority_code), "NONE",
                                            balancing_authority_code)) %>%
  
  # Also turn certain numerical variables into factors
  mutate_at(.vars = vars("sector_code", "balancing_authority_code", "state_code"),
            .funs = factor) 



```



```{r Initial Graphs}
# Graphs Checking out the data
ggplot(data_full, aes(x = log(generation_k_wh), y = log(emissions))) +
  geom_point(aes(color = aggregated_fuel_group)) +
  labs(x = "Generation (KWH)", y = "Tons of CO2 Emissions",
       title = "Power Generated Against CO2 Emissions by Fuel Group (Logged Values)",
      color = "Fuel Group")


ggplot(data_full, aes(y = aggregated_fuel_group, fill = aggregated_fuel_group)) + 
  geom_bar(stat = "count") +
  labs(x = "Total", y = "Fuel Group", 
       caption = "PET = Petroleum, MSW = Municipal solid waste, GEO = Geothermal, GAS = Natural Gas, COAL = Coal") +
  theme_bw() +
  theme(legend.position = "none")

  # Note: Gas is ALL natural gas. MSW is Munincipal Solid Waste. PET is petroleum. 
          # Coal is mostly Bituminous Coal. PET is petroleum, mostly Distillate Fuel Oil
          # GEO is geothermal
  # full table explaining fuel codes here:
    # https://www.eia.gov/electricity/annual/html/epa_a_03.html

```

# Tuning and Training

### Split into Training and Testing Data
```{r training / testing split}
# Set seed before splitting
set.seed(231240)

# Initial 80/20 split for the training/testing data
data_split = data_full %>% initial_split(prop = 0.8)

  # Create training data
train_df = data_split %>% training()

  # Create testing data
test_df = data_split %>% testing()

```

### Create and Tune 4 models on Training data
```{r Creating and Tuning my models}
# ------------------------------ Set up --------------------------------
# Creating Recipe used for multiple models below
train_recipe = train_df %>% recipe(emissions ~ .) %>%
          
                      # update role
                      update_role(plant_code, new_role = "id variable") %>%
  
                      step_dummy(all_nominal()) %>%

  
                      # Mean imputation for numeric predictors
                      step_meanimpute(all_predictors() & all_numeric()) %>% 
  
                      # normalize
                      step_normalize(all_predictors() & all_numeric()) %>%
                      
                      # Remove low-variance
                      step_nzv(all_predictors() & all_numeric()) %>%
  
                      # Remove highly correlated
                      step_corr(all_predictors() & all_numeric()) %>%
  
                      # Remove linearly dependent predictors
                      step_lincomb(all_predictors() & all_numeric()) 
                            







                      
  
                      # Log the energy generation variable
                      #step_log(generation_k_wh)
  
                      # Make interaction variables

  # Meaning of interactions: 
   # sector_group*fuel_code - relationship b/w consumers and types of fuel burnt for energy
   # state_code*cooling_days, state_code*heating_days - how different areas of the country
   # demand energy at different levels depending on temperature
  

  
train_recipe %>% prep() %>% juice()


# setting new seed for CV folds
set.seed(1387219452)
  # Create Cross Validcation folds
train_cv = train_df %>% vfold_cv(v = 5)


```

```{r Models}

# --------------------- Basic linear regression ----------------------

# Fitting the linear regression model to training data
reg_model = 
  linear_reg() %>%
  set_engine("lm")

# Create workflow
workflow_reg = workflow() %>%
  add_model(reg_model) %>%
  add_recipe(train_recipe)

# Train model on validation sets, set metric to root mean square error
cv_reg = workflow_reg %>% 
  fit_resamples(train_cv, metrics = metric_set(rmse))



# ------------------------ Elasticnet --------------------------------

# DEFINE ELASTICNET MODEL
en_model = linear_reg(penalty = tune(), mixture = tune()) %>% 
            set_engine("glmnet")

  # Define workflow
workflow_en = workflow() %>%
  add_model(en_model) %>%
  add_recipe(train_recipe) 

  # Define Penalties
lambdas = 10^seq(from = 5, to = -2, length = 1e2)
alphas = seq(from = 0, to = 1, by = 0.1)

  # Cross Validation Tuning with range of penalites and mixtures
cv_en = workflow_en %>%
  tune_grid(
    train_cv,
    grid = expand_grid(mixture = alphas, penalty = lambdas),
    metrics = metric_set(rmse)
  )

  # Finalize workflow for final model
final_en = workflow_en %>%
  finalize_workflow(select_best(cv_en, 'rmse'))

# Wants penalty = 0.01, mixture = 0.3



# ------------------------------ Random Forest --------------------------------


# One more (Logistic Regression?)

```

### Estimate CV Error
```{r}
#  Estimate your error (with an appropriately chosen metric) using cross validation

# Linear Regression Root Mean Square Error
cv_reg %>% collect_metrics()






# Good place to Create figures!!!!
```

# Prediction

### Test on test data
```{r, include = F}
# -------------------- Predicting with Linear Regression ------------------



```

### Turn in: 
  1. Link to this github page with (commented) analysis code and figures
  2. 1-page "executive summary"
  3. 4-slide summary following paragraphs 2–5 of the executive summary
  4. Evaluation of your group member's contribution (doesn't apply)



